---
title: "Introduction to genreclassifier"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to genreclassifier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(genreclassifier)
```

The genreclassification package makes exploratory data analysis for a classification project easier and more efficient:

-   By cleaning your data as you read it in
-   Counting the proportion of unique factors in the data
-   Visualizing the tidied data with faceted histogram plots

### Data : iris

For the purpose of this demonstration, we will be using the default `iris` dataset from R.

```{r}
data("iris")
dim(iris)
head(iris)
```

Although we are able to easily import this dataset using built-in R functions, for the purpose of classification we will use the `classy_read()` function for its convenient built in functions.

You can see that this is the same dataset, obtained from a Github URL as is often done for reproducibility. 

*Be aware that in this version of the `iris` dataset, the column names are slightly different.*
```{r}
iris_url <- ("https://gist.githubusercontent.com/curran/a08a1080b88344b0c8a7/raw/0e7a9b0a5d22642a06d3d5b9bcbad9890c8ee534/iris.csv")
```

### Read tidy data with classy_read()

Now, we will read it in with `classy_read()` in its simplest form. The dataset is now read from the source as is.

The number of NaNs are summarized.

```{r}
regular_data <- classy_read(iris_url, col = NULL)
dim(regular_data)
head(regular_data)
```

Specify a `col` to convert a column to type `factor`:

```{r}
factored_data <- classy_read(iris_url, "species")
dim(factored_data)
head(factored_data)
```

Optionally provide additional parameters to select the relevant columns for your classification.

```{r}
classy_data <- classy_read(iris_url, "species", species, petal_length:petal_width)
dim(classy_data)
head(classy_data)
```

### Count proportions with count_propotion()

`count_proportion` checks the proportions of unique data values in a column and the proportion of the data that each value makes up.

```{r}
prop <- count_proportion(classy_data, "species", "tidied")
```

### Display your exploratory data with create_faceted_hist_plot()

Two examples below show how you can use `create_faceted_hist_plot()` to create faceted histogram plots for each variable in `row`and align them for EDA.

```{r}
create_faceted_hist_plot(classy_data, "petal_width", "species")
```

```{r}
create_faceted_hist_plot(classy_data, "petal_length", "species")
```















